---
title: "Assignment 4"
---

# Text Analytics Using Quanteda

## Question 1: Quanteda Tutorial

Quanteda workflow \> create a corpus, tokenize the text, remove unnecessary words, build a matrix, then analyze word patterns. Quanteda is useful because it converts large collections of text into structured data that can be compared.

\[Quanteda Tutorial\] (<https://tutorials.quanteda.io>)

## Question 2a - U.S. Presidential Inaugural Speeches

The inaugural speeches can be compared by using word frequencies, common themes, and differences across presidents or parties. Similarities will also appear as themes will be repeats, such as nation, people, freedom, government, while differences can reflect historical context, political party, or the issues emphasized by each administration.

### Set up Environment

```{r}
#| label: setup
#| message: false
#| warning: false

packageneeded <- c(
  "pdftools",
  "tidyverse",
  "quanteda",
  "quanteda.textstats",
  "quanteda.textplots",
  "quanteda.textmodels"
)

new_packages <- packageneeded[
  !packageneeded %in% installed.packages()[, "Package"]
]

if (length(new_packages)) {
  install.packages(new_packages, dependencies = TRUE)
}

library(pdftools)
library(tidyverse)
library(quanteda)
library(quanteda.textstats)
library(quanteda.textplots)
library(quanteda.textmodels)


```

## Question 2b - U.S. Department of Defense Reports

The DOD reports were analyzed by reading the PDF files, converting them into text, and creating a quanteda corpus. After tokenizing and cleaning the text, the reports were transformed into a document feature matrix. Allowing the analysis to identify the most common terms, track key defense relation words over time, and compare language across presidential administrations.

### Analyze U.S. Department of Defense Reports

```{r}

PATH_DATA <- "/Users/patrickheldebrandt/Desktop/EPPS6323/USDOD"

pdf_files <- list.files(
  PATH_DATA,
  pattern = "\\.pdf$",
  full.names = TRUE,
  ignore.case = TRUE
)

cat("Found", length(pdf_files), "PDF files\n")



read_pdf_safe <- function(filepath) {
  tryCatch({
    pages <- pdftools::pdf_text(filepath)
    text <- paste(pages, collapse = "\n")
    return(text)
  }, error = function(e) {
    warning("Failed to read: ", basename(filepath))
    return(NA_character_)
  })
}

DODq <- tibble(
  doc_id = basename(pdf_files),
  text = map_chr(pdf_files, read_pdf_safe)
)

DODq <- DODq %>%
  filter(!is.na(text))

DODq$year <- as.numeric(str_extract(DODq$doc_id, "\\d{4}"))

DODq$doc_id <- paste0("USDOD_", DODq$year, ".pdf")

DODq <- arrange(DODq, year)

cat("Successfully loaded", nrow(DODq), "reports\n")



corp_DOD <- corpus(DODq, text_field = "text")

corp_DOD$year <- DODq$year


corp_DOD$president <- case_when(
  corp_DOD$year <= 2000 ~ "Clinton",
  corp_DOD$year >= 2001 & corp_DOD$year <= 2008 ~ "Bush",
  corp_DOD$year >= 2009 & corp_DOD$year <= 2016 ~ "Obama",
  corp_DOD$year >= 2017 & corp_DOD$year <= 2020 ~ "Trump",
  corp_DOD$year >= 2021 & corp_DOD$year <= 2024 ~ "Biden",
  corp_DOD$year >= 2025 ~ "Trump II",
  TRUE ~ NA_character_
)

summary(corp_DOD, 5)
table(corp_DOD$president)


toks_DOD <- tokens(
  corp_DOD,
  remove_punct = TRUE,
  remove_numbers = TRUE
) %>%
  tokens_remove(stopwords("english")) %>%
  tokens_remove(c(
    "page", "figure", "table", "appendix", "chapter",
    "https", "www", "pdf", "gov"
  ))

dfmat_DOD <- dfm(toks_DOD)

cat("DFM dimensions:", dim(dfmat_DOD), "\n")
cat("Documents:", ndoc(dfmat_DOD), "| Features:", nfeat(dfmat_DOD), "\n")

topfeat <- topfeatures(dfmat_DOD, 30)

print(topfeat)

data.frame(
  feature = names(topfeat),
  freq = topfeat
) %>%
  ggplot(aes(x = reorder(feature, freq), y = freq)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Top 30 Features across DOD China Reports",
    x = NULL,
    y = "Frequency"
  ) +
  theme_bw(base_size = 14)


textplot_wordcloud(
  dfmat_DOD,
  max_words = 100
)


dfmat_year <- dfm_group(
  dfmat_DOD,
  groups = corp_DOD$year
)

key_terms <- c(
  "taiwan", "pla", "nuclear", "cyber", "missile",
  "threat", "aircraft", "carrier", "space", "ai"
)

term_trends <- convert(dfmat_year, to = "data.frame") %>%
  select(doc_id, any_of(key_terms)) %>%
  mutate(year = as.numeric(str_extract(doc_id, "\\d{4}"))) %>%
  pivot_longer(
    cols = -c(doc_id, year),
    names_to = "term",
    values_to = "count"
  )

doc_lengths <- ntoken(dfmat_year)

term_trends <- term_trends %>%
  left_join(
    tibble(doc_id = names(doc_lengths), total = doc_lengths),
    by = "doc_id"
  ) %>%
  mutate(rate = count / total * 10000)

ggplot(term_trends, aes(x = year, y = rate)) +
  geom_line() +
  geom_point() +
  facet_wrap(~term, scales = "free_y") +
  labs(
    title = "Key Term Frequency in DOD China Reports",
    subtitle = "Rate per 10,000 words",
    x = "Year",
    y = "Rate"
  ) +
  theme_bw(base_size = 14)

```

## What is Wordfish

Wordfish is able to estimate the latent position of documents based on word usage. Documents that use similar vocabulary are placed closer together, while documents with different language patterns are placed further apart. We use Wordfish to help compare how the language of DOD reports change over time.

```{r}

tstat_key <- textstat_keyness(
  dfmat_DOD,
  target = corp_DOD$president %in% c("Trump", "Trump II"),
  measure = "lr"
)

textplot_keyness(tstat_key, n = 20) +
  labs(
    title = "Keyness: Trump-era vs. Other Administrations",
    subtitle = "Words more associated with Trump-era DOD reports"
  ) +
  theme_bw(base_size = 14)


dfmat_wf <- dfm_trim(
  dfmat_DOD,
  min_termfreq = 10,
  max_docfreq = 0.95,
  docfreq_type = "prop"
)

anchor_early <- which(docnames(dfmat_wf) == "USDOD_2000.pdf")
anchor_late <- which(docnames(dfmat_wf) == "USDOD_2023.pdf")

cat(
  "Anchor documents:",
  docnames(dfmat_wf)[anchor_early],
  "and",
  docnames(dfmat_wf)[anchor_late],
  "\n"
)

tmod_wf <- textmodel_wordfish(
  dfmat_wf,
  dir = c(anchor_early, anchor_late)
)

summary(tmod_wf)

```

## Question 3a - Run Wordfish on DOD reports

First we trim the document feature matrix to remove very rare and overly common terms. To set the direction of the scale we select anchor documents, this allows us to use the Wordfish model to estimate a theta value for each report, which represents the report's position on the latent text scale.

### The Compare Positions Using Scaling Methods

```{r}


doc_pos <- data.frame(
  document = docnames(dfmat_wf),
  theta = tmod_wf$theta,
  year = as.numeric(str_extract(docnames(dfmat_wf), "\\d{4}")),
  president = dfmat_wf$president,
  stringsAsFactors = FALSE
) %>%
  arrange(year)

print(doc_pos)

ggplot(doc_pos, aes(x = year, y = theta)) +
  geom_line(linetype = "dashed") +
  geom_point(aes(color = president), size = 3) +
  geom_text(aes(label = year), vjust = -1, size = 3) +
  geom_smooth(method = "loess", se = TRUE) +
  labs(
    title = "DOD China Report Positioning by Year",
    subtitle = "Estimated Wordfish latent position",
    x = "Year",
    y = "Theta",
    color = "President"
  ) +
  theme_bw(base_size = 14)


textplot_scale1d(tmod_wf)

textplot_scale1d(
  tmod_wf,
  groups = dfmat_wf$president
)

textplot_scale1d(
  tmod_wf,
  margin = "features",
  highlighted = c(
    "taiwan", "pla", "nuclear", "missile",
    "cyber", "space", "carrier", "ai"
  )
)

tmod_ca <- textmodel_ca(dfmat_wf)

dat_ca <- data.frame(
  dim1 = tmod_ca$rowcoord[, 1],
  dim2 = tmod_ca$rowcoord[, 2],
  document = rownames(tmod_ca$rowcoord),
  stringsAsFactors = FALSE
)

dat_ca$year <- as.numeric(str_extract(dat_ca$document, "\\d{4}"))
dat_ca$president <- dfmat_wf$president

ggplot(dat_ca, aes(x = dim1, y = dim2, color = president)) +
  geom_point(size = 3) +
  geom_text(aes(label = year), vjust = -1, size = 3) +
  labs(
    title = "Correspondence Analysis: DOD China Reports",
    x = "Dimension 1",
    y = "Dimension 2",
    color = "President"
  ) +
  theme_minimal(base_size = 14)



```

## Question 4

\*\*Wordfish\*\* is a one dimensional scaling model that estimate the latent position of each document based on patterns in word usage. This makes it useful for identifying broad shifts in strategic language or emphasis across years and presidential administrations.

\*\*Correspondence\*\* analysis provides a multidimensional comparison of reports, as opposed to placing documents along a single scale, correspondence analysis can display documents and terms across multiple dimensions making it easier to identify clusters and relationship among reports.

## Final Summary

```{r}

cat("\n========== ANALYSIS SUMMARY ==========\n")
cat("Reports analyzed:", ndoc(corp_DOD), "\n")
cat("Year range:", min(DODq$year), "-", max(DODq$year), "\n")
cat("Presidents:", paste(unique(corp_DOD$president), collapse = ", "), "\n")
cat("DFM dimensions:", dim(dfmat_DOD), "\n")
cat("Wordfish theta range:", round(range(tmod_wf$theta), 3), "\n")
cat("=======================================\n")
```
